---
title: "Understand the Impact of Demographic and Labor Market Trends on Economic Growth in East Asia"
subtitle: "Exploring the Role of Aging, Unemployment, and Workforce Participation in GDP per Capita from 1990 to 2023"
author: 
  - Ruiyang Wang
thanks: "Code and data are available at: [https://github.com/Ruiyang-Wang/STA304-Final-East-Asia-GDP-Analysis.git](https://github.com/Ruiyang-Wang/STA304-Final-East-Asia-GDP-Analysis.git)."
date: today
date-format: long
abstract: "This paper examines the impact of demographic changes and labor market trends on economic growth in East Asia from 1990 to 2023. Analyzing data such as GDP per capita, unemployment rates, hours worked, and labor force participation, the study shows that while East Asia has experienced overall economic growth, the effects of an aging population and a shrinking working-age group have begun to slow this progress. The findings underscore that countries such as Japan and China are facing increasing challenges from demographic changes that could reduce their economic potential. This study is important because it highlights the importance of considering demographic trends when formulating future economic policies and projections for the region."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
#| warning: false
#| message: false

library(here)
library(tidyverse)
library(ggplot2)
library(reshape2)
library(car)
library(lmtest)
library(modelsummary)
east_asia_data <- read.csv(here("data/02-analysis_data/analysis_data.csv"))
```

# Introduction

Over the past few decades, East Asia has undergone a major economic
transformation, driven by rapid industrialization, urbanization, and
demographic changes. However, as these countries face the challenges of
aging populations, declining birth rates and stagnant economic growth,
it is significant to understand how these factors interact with key
economic indicators such as GDP per capita. This paper examines the
relationship between demographic trends, labor market dynamics, and
economic performance in East Asia from 1990 to 2023.

Despite East Asia showed its historical economic success during later
20th century and early 21st century, there are problems emerging recent
years in demographic changes, especially related to aging population and
shrinking working-age groups. By applying a linear regression model to
the dataset by @economist_gdp_dataset, this paper aims to fill this gap
by assessing the role of population aging, labor force participation
rates, and other predictors in shaping East Asia's economic growth.

The analysis found that while GDP per capita in East Asia has generally
grown over time, the contribution of demographic factors such as the
share of the working-age population and the labor force participation
rate has varied. In addition, models with interaction terms and multiple
predictors provide explanations into the diminishing impact of the
demographic dividend, especially in countries such as Japan and China.
These findings highlight the importance of considering demographic
trends in economic forecasts and decision-making.

## Estimand

The main estimand of this paper is to estimate the relationship between
population and labor market factors and GDP per capita in East Asia
during 1990-2023. This paper aims to measure how variables such as
population aging, labor force participation, unemployment, and hours
worked per capita affect economic growth in the region. This includes
assessing the role of time (years) in GDP growth trends, examining how
changes in the working-age and elderly populations affect GDP per
capita, and exploring how the unemployment rate and labor force
participation rate affect economic performance. In addition, the study
looks at the interaction between population aging and labor force
participation to understand how these factors work together to influence
economic outcomes. Ultimately, the paper seeks to quantify how
demographic changes and labor market dynamics contribute to economic
growth in East Asia over the past few decades.

## Structure of the Paper

@sec-data describes the dataset and key variables by presenting graphs.
@sec-regression presents the simple linear regression (SLR), multiple
linear regression (MLR) models used to explore the effect of sample size
on support percentage. @sec-result presents the analytical results for
dataset graphs and regression models. @sec-disc discusses further
discussions based on paper discoveries.

\newpage

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR] to analyze the
dataset from @economist_gdp_dataset. This dataset provides economic
indicators, including GDP per capita, unemployment rates, and hours
worked per population for countries worldwide.

@sec-rawpreview provides detailed economic indicators across multiple
countries and years, with a particular focus on GDP per capita, labor
market participation, and productivity measures. The dataset is
accessible through The Economist's GitHub repository, which hosts the
raw data and provides transparent access to economic figures from 1990
to 2023 for a variety of countries worldwide.

## Data Measurement

Measurements in the dataset are standardized to ensure consistency and comparability across countries and years. GDP per capita is calculated by dividing a country's gross domestic product (GDP) by its total population. To account for inflation and ensure that values are consistent across time periods, GDP data is adjusted to real GDP for more accurate cross-country comparisons. The unemployment rate reflects the percentage of the labor force that is actively looking for work. The number of hours worked per person is calculated by dividing the total number of hours worked by the entire population to arrive at the sum of labor productivity. This indicator is particularly useful in analyzing the effects of labor intensity and productivity on economic performance. Finally, demographic data, including age-specific statistics, are derived from national census data, ensuring consistency across countries. Age-related variables, such as the proportion of people aged 65 and over, are adjusted according to each country's population reporting standards to ensure comparability across time periods.

## Data Consideration

To ensure the consistency, most missing values were imputed by utilizing
R functions in combination of `dplyr` and `lahman` packages
[@citeLahman]. Countries outside of East Asia were excluded from the
analysis. However,in order to make sure the intergrity of whole dataset,
some NA values are not eliminated, since they only occur once or twice
in a row, and all missing values has a pattern to follow, which can be
avoided in later graphing and analyzing.

## Methodology

In order to analyze the relationship between economic indicators and
population trends, we use linear regression model. The method involves
fitting simple and multiple regression models to estimate the effects of
various predictors, such as the labor force participation rate,
population aging, unemployment rate, and hours worked, on GDP per
capita. Interaction terms are included in the model to assess how
certain demographic factors, such as population aging and labor force
participation, interact and collectively affect economic growth. The
analysis was performed using the `rstanarm` package in R [@citeR], using
default priors for Bayesian regression models to ensure robust results.
The statistical techniques used in this paper provide a detailed
understanding of the dynamics between demographic trends and economic
performance over the past few decades.

## Data Visualization

### GDP per Capita {#sec-GDPc}

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: GDPc
#| fig.cap: ""
#| fig-pos: "H"
# Load the saved RDS file for GDP Per Capita Trends
gdp_trends <- readRDS("~/STA304_Final/models/gdp_trends.rds")
gdp_trends

```

The per capita GDP trends shown in [@sec-GDPc] show significant
differences between East Asian countries from 1990 to 2023. China's
rapid economic growth reflects the economic transformation driven by
industrialization, urbanization and integration into the global market.
In contrast, Japan's GDP per capita remained stable, indicating a mature
and developed economy with limited growth potential. The per capita GDP
of the Chinese Macao SAR has consistently exceeded that of other
countries due to its specialized economy, which relies heavily on
tourism and gambling industries (@greenwood2016reinventing). However,
countries such as Mongolia and North Korea have significantly lower GDP
per capita, highlighting the challenges they face in terms of
development and structural economies.

### Unemployment Rate {#sec-unemployr}

```{r}
#| label: unemployr
#| echo: false
#| warning: false
#| message: false
#| fig.cap: "Unemployment Rate Trends in East Asia (1990-2023)"
#| fig-pos: "H"
# Load the saved RDS file for Unemployment Rate Trends
unemployment_trends <- readRDS("~/STA304_Final/models/unemployment_trends.rds")
unemployment_trends


```

The unemployment rate trend shown in [@sec-unemployr] displays
significant differences in East Asia. Mongolia's unemployment rate
fluctuated the most, likely due to volatility in its resource-dependent
economy(@MongoliaDependent). In contrast, Japan and China's Hong Kong
Special Administrative Region have maintained relatively stable and low
unemployment rates, emphasizing their stable economies and efficient
labor markets. China's unemployment rate is fluctuating, displays
instability.

### Hours Worked Per Population {#sec-hr_work}

```{r}
#| label: hr_work
#| echo: false
#| warning: false
#| message: false
#| fig.cap: "Hours Worked Per Population in East Asia (1990-2023)"
#| fig-pos: "H"

# Load the saved RDS file for Hours Worked Per Population Trends
hours_worked_trends <- readRDS("~/STA304_Final/models/hours_worked_trends.rds")
hours_worked_trends


```

The unemployment rate trend shown in [@sec-unemployr] displays
significant differences in East Asia. Mongolia's unemployment rate
fluctuated the most, likely due to volatility in its resource-dependent
economy(@MongoliaDependent). In contrast, Japan and China's Hong Kong
Special Administrative Region have maintained relatively stable and low
unemployment rates, emphasizing their stable economies and efficient
labor markets. China's unemployment rate is fluctuating, displays
instability.

## Predictor variables

### Population Aged 65 and Older {#sec-older_age}

```{r}
#| label: older_age
#| echo: false
#| warning: false
#| message: false
#| fig.cap: "Population Aged 65+ vs GDP Per Capita in East Asia (1990-2023)"
#| fig-pos: "H"
# Load the saved RDS file for Population Aged 65+ vs GDP Per Capita
pop_over_65_vs_gdp <- readRDS("~/STA304_Final/models/pop_over_65_vs_gdp.rds")
pop_over_65_vs_gdp

```

The unemployment rate trend shown in [@sec-unemployr] displays
significant differences in East Asia. Mongolia's unemployment rate
fluctuated the most, likely due to volatility in its resource-dependent
economy(@MongoliaDependent). In contrast, Japan and China's Hong Kong
Special Administrative Region have maintained relatively stable and low
unemployment rates, emphasizing their stable economies and efficient
labor markets. China's unemployment rate is fluctuating, displays
instability.

### Youth Population {#sec-youth_age}

```{r}
#| label: youth_age
#| echo: false
#| warning: false
#| message: false
#| fig.cap: "Youth Population vs Unemployment Rate in East Asia (1990-2023)"
#| fig-pos: "H"
# Load the saved RDS file for Youth Population vs Unemployment Rate
youth_vs_unemployment <- readRDS("~/STA304_Final/models/youth_vs_unemployment.rds")
youth_vs_unemployment


```

The unemployment rate trend shown in [@sec-unemployr] displays
significant differences in East Asia. Mongolia's unemployment rate
fluctuated the most, likely due to volatility in its resource-dependent
economy(@MongoliaDependent). In contrast, Japan and China's Hong Kong
Special Administrative Region have maintained relatively stable and low
unemployment rates, emphasizing their stable economies and efficient
labor markets. China's unemployment rate is fluctuating, displays
instability.

### Labor Force Participation {#sec-labor_participation}

```{r}
#| label: labor_participation
#| echo: false
#| warning: false
#| message: false
#| fig.cap: "Labor Force Participation vs GDP Per Capita in East Asia (1990-2023)"
#| fig-pos: "H"
# Load the saved RDS file for Labor Force Participation vs GDP Per Capita (China)
china_plot <- readRDS("~/STA304_Final/models/china_plot.rds")
china_plot

# Load the saved RDS file for Labor Force Participation vs GDP Per Capita (Other East Asian Countries)
other_countries_plot <- readRDS("~/STA304_Final/models/other_countries_plot.rds")
other_countries_plot

```

As can be seen from the two labor participation rate charts shown in
[@sec-labor_participation], China's labor force is huge, and the labor
participation rate presents a unique pattern. While GDP per capita
increases as labor force participation increases, the relationship is
non-linear, indicating diminishing returns or other influencing factors
such as productivity and industrial relocation. For other East Asian
countries, smaller workforces exhibit more of a cluster pattern, with
countries like Japan and China's Hong Kong Special Administrative Region
having high GDP per capita despite a smaller workforce, highlighting the
role of high productivity and economic efficiency.

# Linear Regression Analysis {#sec-regression}

We did our Linear regression analysis based on the `tidyverse` package.
The goal of our regression analysis is divided into three steps. First,
we aim to capture the general trends in GDP per capita over time.
Second, we explore how interaction terms affect GDP per capita. Third,
we want to find out how additional predictors, including unemployment,
labor force participation, aging population, and working hours,
contribute to GDP per capita.

\newpage

## Model 1: Simple Linear Regression {#sec-SLRmodel}

Define ( y_i ) as the GDP per capita for country ( i ) in year ( t ).
Let ( \text{Year}\_i ) represent the year. The model is specified as:

\[ \begin{align}
y_i &\sim \text{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \cdot \text{Year}_i
\end{align} \]

We run the model in R [@citeR] using the `rstanarm` package of
@rstanarm. We use the default priors from `rstanarm`.We fit this model
in R using the `lm()` function, and the result is shown above.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: modelSLR
#| fig-pos: "H"
# Simple Linear Regression
simple_model <- lm(gdp_over_pop ~ year, data = east_asia_data)
modelsummary::modelsummary(simple_model)
```

This linear regression model validation is located in Appendix at
[@sec-DSLR].

\newpage

## Model 2: Interaction Terms {#sec-INTmodel}

To explore whether the effects of one predictor depend on another, we
include interaction terms in the model. Specifically, we examine how the
relationship between aging population percentage ( \text{AgingPop}\_i )
and GDP per capita is influenced by labor force participation (
\text{LaborForce}\_i ).

Define ( y_i ) as the GDP per capita for country ( i ) in year ( t ):

\[ \begin{align}
y_i &\sim \text{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \cdot \text{AgingPop}_i + \beta_2 \cdot \text{LaborForce}_i + \beta_3 \cdot (\text{AgingPop}_i \cdot \text{LaborForce}_i)
\end{align} \]

We run the model in R [@citeR]. We fit this model in R using the `lm()`
function and result is shown above.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: model_INT
#| fig-pos: "H"
# Multiple Linear Regression with Interaction Terms
interaction_model <- lm(gdp_over_pop ~ pop_15_to_64 * labor_force, data = east_asia_data)
modelsummary::modelsummary(interaction_model)
```

This linear regression model validations are located in Appendix at
[@sec-DINT].

\newpage

## Model 3: Multiple Linear Regression {#sec-MLRmodel}

We extend our analysis to include additional predictors: unemployment
rate ( \text{Unemployment}\_i ), labor force participation (
\text{LaborForce}\_i ), aging population percentage ( \text{AgingPop}\_i
), and hours worked ( \text{HoursWorked}\_i ). Define ( y_i ) as the GDP
per capita for country ( i ) in year ( t ):

\[ \begin{align}
y_i &\sim \text{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \cdot \text{Unemployment}_i + \beta_2 \cdot \text{LaborForce}_i + \beta_3 \cdot \text{AgingPop}_i + \beta_4 \cdot \text{HoursWorked}_i
\end{align} \]

We run the model in R [@citeR]. We fit this model in R using the `lm()`
function and result is shown above.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: modelMLR
#| fig-pos: "H"
# Multiple Linear Regression
multiple_model <- lm(gdp_over_pop ~ unemployment_r + labor_force + pop_over_65 + hours_worked_over_pop_combined, 
                     data = east_asia_data)
modelsummary::modelsummary(multiple_model)
```

This linear regression model validations are located in Appendix at
[@sec-DMLR].

\newpage

# Result {#sec-result}

## GDP per capita with time period

The results of a simple linear regression analysis in [@sec-SLRmodel]
provide the relationship between GDP per capita and time in East Asia.

The slope coefficient of the annual variable is estimated to be 811.76,
indicating that GDP per capita increases by about 811.76 units per year
on average. This positive trend is statistically significant (p \<
0.001), indicating that time is the key factor driving East Asian
economic growth over the period analyzed (1990-2023). This result is in
line with a broader understanding of the region's economic development,
which is characterized by continued growth and industrialization.

The model explains about 15.3% of the change in GDP per capita (R
squared = 0.153), with an adjusted R squared of 0.1487, suggesting that
the year variable reflects only a small part of the change in East Asian
economies. This suggests that other factors, such as labor force
dynamics, demographic changes, and policy interventions, play a key role
in explaining trends in GDP per capita.

## GDP per capita with Working-Age Population and Labor Force Participation

The interaction model in [@sec-INTmodel] introduces a key dimension to the
analysis by looking at how the relationship between the share of the
population aged 15-64 and GDP per capita depends on the labor force
participation rate.

The coefficient for the percentage of the population aged 15 to 64 was
estimated at 249,600, indicating a strong positive correlation with GDP
per capita. Every one unit increase in this percentage increases GDP per
capita by about 249,600 units. This result is highly statistically
significant (p\<0.001), highlighting the importance of this demographic
in driving economic performance.

The positive coefficient of the labor force participation rate is
1.191×10\^-3, which means that a higher labor force participation rate
contributes positively to GDP per capita. This result (p=0.000728)
reflects the role of labor force participation in economic productivity.

The interaction term between the proportion of the population aged 15 to
64 and the labor force participation rate is a negative coefficient
(-1.742 ×10\^−3), indicating that the positive impact of the working-age
population on GDP per capita decreases as the labor force participation
rate increases. This finding (p=0.000499) highlights the complexity of
the relationship between demographic composition and economic outcomes.
It may reflect diminishing returns to labor, or the challenge of
integrating a growing workforce into productive economic activity.

The overall model explains 47.51% of the change in GDP per capita (R
squared =0.4751), with an adjusted R squared of 0.452. Although this is
a lower explanatory power than a multiple linear regression model, the
inclusion of interaction terms gives some explanations to the
interaction between population and labor factors.

## GDP per capita with multiple factors

The multiple linear regression model in [@sec-MLRmodel] is introduced to
increase the overall explanatory power of the model.

The unemployment rate coefficient is estimated at -928.4, indicating
that every unit increase in the unemployment rate reduces GDP per capita
by about 928.4 units. However, with a P-value of 0.0554, this predictor
is not statistically significant at the 5% significance level,
suggesting that its effect on GDP per capita may be limited or
influenced by other factors in the model.

The coefficient of the labor force participation rate is -3.366 ×10\^-5,
indicating that it is negatively correlated with GDP per capita. This
suggests that an increase in the labor force could dilute GDP per capita
without a corresponding increase in productivity or economic capacity.
This prediction is highly significant (p\<0.001).

The proportion of population aging has a significant positive impact on
per capita GDP, and its coefficient is 2136.0. This means that for every
unit increase in the proportion of people over 65, GDP per capita
increases by about 2,136 units. This finding reflects the economic
structure of developed East Asian countries, where aging populations
tend to be associated with higher levels of productivity and income.

Finally, hours worked per capita were also significantly positively
correlated with GDP per capita (β=7.463, p\<0.001), suggesting that
higher average hours worked contributed to an increase in GDP per
capita, which may reflect labor productivity and economic effort.

The overall model explains about 81.5% of the variation in GDP per
capita (R squared =0.8153), with an adjusted R squared of 0.8114. This
shows a substantial improvement in explanatory power compared to simple
linear regression models.

\newpage

# Discussion {#sec-disc}

## Choice of full model or reduced model

In the multiple regression analysis, the summary table shows that the
unemployment rate is not as influential as other predictors, an ANOVA
test is required to compare the explanatory power between full model and
reduced model.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: ANOVA
#| fig-pos: "H"
full_model<- lm(gdp_over_pop ~ unemployment_r + labor_force + pop_over_65 + hours_worked_over_pop_combined, 
                     data = east_asia_data)
reduced_model <- lm(gdp_over_pop ~ labor_force + pop_over_65 + hours_worked_over_pop_combined, 
                     data = east_asia_data)
anova(full_model, reduced_model)
```

The ANOVA test results show that when we compare the full model (which
includes the predictor `unemployment_r`) to the reduced model (which does
not include `unemployment_r`), the p-value is 0.05538. This indicates that
the difference in model fit is not statistically significant at the
usual 0.05 threshold. In other words, removing the unemployment_r
predictor from the full model does not lead to a substantial loss in
explanatory power, which implies we should keep the predictor for the
overall data integrality.

## Diminishing Impact of the Demographic Dividend

As Mason says in his book "Demography and the Asian
Economic Miracle", an increase in the working-age population ratio would
lead to an increase in income levels (@mason1997population). That rapid increase in the
working-age population due to a low birth rate would produce a
"demographic dividend." However, demographic dividend occur less as less
as time goes on.

In recent years, major East Asian countries discussed in this paper have
faced economic slowdowns. Countries like Japan, South Korea, and China,
which once benefited from a growing working-age population, are now
encountering challenges related to aging populations and declining birth
rates.

Japan, has suffered severe economic stagnation in recent decades
(@hoshi2004japan). Although Japan was one of the first countries to
experience a "demographic dividend" during the post-World War II
economic boom, its working-age population is shrinking due to low
fertility rates and an aging society. According to Takeo Hoshi, Japan's
GDP growth has slowed, and the country faces increasing social welfare
costs for its elderly population. Against this backdrop, the
"demographic dividend" that fueled Japan's economic rise appears to have
largely dissipated, replaced by challenges such as labor shortages and
falling consumer demand.

In China, the one-child policy (which lasted from 1979 to 2015) reduced
the birth rate, resulting in a growing elderly population. While China
benefited from a demographic dividend during its rapid industrialization
and economic growth shown in [@sec-GDPc], the country is likely to face
a "demographic crisis" in the future since its working-age population
begins to shrink. Despite the effect of COVID-19, part of China's
slowdown can be attributed to a shrinking workforce and the growing
fiscal burden in order to feed the aging population.

In conclusion, while the demographic dividend once provided a huge boost
to East Asia's economies, the region's major economies are now facing
the negative effects of aging populations and declining birth rates.
Over time, the demographic shift from a young, expanding workforce to an
aging, shrinking population is becoming more evident in the economic
performance of these countries, signaling that the era of demographic
dividends may be coming to an end, leaving an significant macroeconomic
problem to the government.

## Impact of Model Complexity on R-Squared

In this paper, we observe that both simple linear regression and linear
regression with interaction terms have R-squared values below 0.5,
suggesting that these models can only explain less than half of the
variance in the outcome variables, which is, GDP per capita. This
suggests that either the model specification is missing important
predictors, or the relationship between the variables is more complex
than what is already captured. A multiple linear regression model,
however, with four variables in this paper, showed a relatively higher
R-squared value of more than 0.8, indicating a much better fit to the
data because including additional variables help explain a large portion
of the variance. While a higher R-square generally indicates a stronger
model, it also emphasizes the importance of selecting relevant
variables; Too many predictors can lead to overfitting, with models
capturing noise rather than the true underlying pattern.

## Weaknesses

Although regression models provide general explanations, they still rely on
certain assumptions that may not fully reflect the complexity of
real-world economic systems. For example, linear regression assumes a
constant relationship between the predictor and the outcome, but this
relationship may not hold in the case of nonlinear interactions or
changes in structure over time. While some terms for interactions are
included to address this problem, the model may not capture other
nonlinear effects.

This paper focuses on a set of demographic and labor market predictors,
but other factors—such as technological advancement, international
trade, government policies, and institutional factors—were not included
in the models. These variables may also significantly affect GDP per
capita and could provide a more complete understanding of the economic
dynamics in East Asia.

The focus on East Asia restricts the analysis to a single region. While
this allows for in-depth exploration of economic trends in this area, it
also means that the findings cannot be generalized to other parts of the
world. A broader analysis that includes countries from different regions
would provide more context and might highlight whether the patterns
observed in East Asia are unique to this region or shared globally.

## Next Steps

Future research could test nonlinear models to better capture the
complexity of the relationship between predictors and GDP per capita.
Methods such as polynomial regression, decision trees, or machine
learning algorithms such as random forests can help model more complex
patterns that current linear models cannot capture. In addition,
extending the model to include other economic and social factors, such
as technological innovation and education levels, would provide a more
realistic view of the drivers of GDP growth. Finally, including
countries in regions outside East Asia, such as Southeast Asia, Europe,
or North America, can help determine whether the patterns found in this
study are specific to East Asia or part of a broader global trend.

\newpage

\appendix

# Appendix - Raw Data {#sec-rawpreview}

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: rawpreview
#| fig-pos: "H"
#| fig-cap: "Preview of the raw data"

raw_data<- read.csv(here("data/01-raw_data/raw_data.csv"))
# Extract the first 5 rows and all 15 columns
data_first_5_rows <- raw_data[7762:7766, 1:15]
# Split the data into 3 parts (5 variables per part)
part1 <- data_first_5_rows[, 1:5]
part2 <- data_first_5_rows[, 6:10]
part3 <- data_first_5_rows[, 11:15]

# Display the table
print(part1)
print(part2)
print(part3)


```
This shows 5 rows in raw data, give an overall preview to the raw data.

# Appendix - Data Validation

## Simple Linear Regression Validation {#sec-DSLR}

To assess the simple regression model, we examine residuals for
linearity, homoscedasticity, and normality,and uncorrelated errors.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: assumpSLR
#| fig-pos: "H"
# Residual Diagnostics for the Simple Linear Regression Model
par(mfrow = c(2, 2))
plot(simple_model)
```

The residual diagnosis shows that while the model is statistically
significant, the variability of the residual points to
heteroscedasticity, while the unexplained variance suggests missing
variables. It can be observed that there are slightly discrete patterns
in the tail of the QQ graph, which does not affect to overall accuarcy.

## Multiple Linear Regression {#sec-DMLR}

To evaluate the assumptions of the MLR model, we examine residuals for
linearity, homoscedasticity, and normality,and uncorrelated errors.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: assumpMLR
#| fig-pos: "H"
# Residual Diagnostics for the Multiple Linear Regression Model
par(mfrow = c(2, 2))
plot(multiple_model)

```

Multicollinearity occurs when predictors are highly correlated with each
other, potentially inflating standard errors. We use the Variance
Inflation Factor (VIF) to assess multicollinearity:

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: VIF_MLR
#| fig-pos: "H"
# Check for multicollinearity using VIF
vif(multiple_model)
```

The residuals should be independent. We test this assumption using the
Durbin-Watson test:

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: DW_MLR
#| fig-pos: "H"
# Test for independence of errors
dwtest(multiple_model)
```

## Multiple Linear Regression (Interaction term included) {#sec-DINT}

To evaluate the assumptions of the model, we examine residuals for
linearity, homoscedasticity, and normality,and uncorrelated errors.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: INT_ASS
#| fig-pos: "H"
# Residual Diagnostics for the Interaction Model
par(mfrow = c(2, 2))
plot(interaction_model)
```

Interaction terms can introduce multicollinearity. As a result, we check
Variance Inflation Factors (VIFs) to ensure predictors remain
interpretable:

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: INT_VIF
#| fig-pos: "H"
# Check for multicollinearity using VIF
vif(interaction_model)
```

We test for autocorrelation in residuals using the Durbin-Watson test:

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: INT_DW
#| fig-pos: "H"
# Test for independence of errors
dwtest(interaction_model)
```

# Appendix - Simulated Data Creation and Its Relationship to Surveys, Sampling, and Observational Data

In this paper, we employ simulated data to model a set of socioeconomic
indicators for countries in East Asia, such as population size, labor
force, GDP, unemployment rates, and age demographics. These variables
mirror the kind of information that might be gathered in national
surveys or from observational data sources. The advantage of using
simulated data in this context is that it allows for controlled
experimentation, where we can test different hypotheses and analytical
techniques without being constrained by the limitations or biases
inherent in real-world datasets.

The simulated data used in this study was created using a random
sampling approach, with each variable being generated according to
predefined distributions or value ranges that approximate real-world
scenarios. For example, in the simulation, the variable representing
population size (`pop`) was generated using a uniform distribution,
where values range from 500,000 to 1.5 billion, reflecting the large
population sizes found in East Asia. Similarly, other variables such as
labor force (`labor_force`), GDP (`gdp`), and unemployment rate
(`unemployment_r`) were modeled using random distributions within
realistic bounds.

The simulation mimics how real-world data might be gathered in surveys
or observational studies. For example, the country variable is sampled
from a predefined list of countries in East Asia, reflecting the target
population of the study. This approach is analogous to stratified or
random sampling in surveys, where the sample is drawn from a specific
population to ensure representativeness. Additionally, the inclusion of
variables such as GDP per capita and hours worked is reflective of the
kind of socioeconomic data often collected in national surveys or labor
force surveys.

\newpage

# References
